---
title: "Draft Proposal"
output:
  html_document:
    code_folding: show

---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE,
                      message = FALSE,
                      options(scipen=999))
```


#### 2020-10-26
### Lika Mikhelashvili, Kelly Ford, Sarah Bingham
**“Hello Statisticians”**

```{r}
library(ipumsr)
library(ggplot2)
library(moderndive)
ddi <- read_ipums_ddi("nhis_00001.xml")
data <- read_ipums_micro(ddi)
```

Members:

* Lika Mikhelashvili  
* Kelly Ford  
* Sarah Bingham  

## Project Title 
Health Factors Related to Sleep

## Purpose and Hypotheses
This project focuses on self-reported areas of mental and physical health among 72,831 randomly selected people living in the United States during the year of 2018. These spheres of health include whether or not one takes depression and/or sleep medication, hours of sleep per day, and amount of alcohol consumed per year. The goal of this project is to compare these variables to the frequency of computer use in order to more clearly visualize the effect screen time on health. This project is very relevant to life during the pandemic, where we depend on technology and computer use more than ever. With this increased screen use, however, we may see a decrease in mental and physical health based on limited in-person social interaction and necessary self-isolation from others. Our hope is that this project gives some insight into the health ramifications associated with computer usage within a large economically and racially diverse sample. 

The **primary hypothesis** is that screen time per week, depression medication, or sleep medication will have a significant effect on self-reported usual hours of sleep per day. 

## Data  
Our study utilizes data made available online to the public from IPUMS-NHIS (National Health Interview Survey). This data set contains information from 2018, the most recent year available. Our variables include age, race, earnings, weight, alcohol use, sleep habits, depression levels and medications, and frequency of computer use.

## Population  
IPUMS-NHIS data is collected annually of approximately 100,000 individuals throughout 45,000 households in the United States. Each observation represents a separate individual, though this does indicate some individuals cohabitate. Each household has its own serial ID to distinguish which data come from which household. As cohabitation does not violate independence, we will consider all of the data points in our analysis. We will analyze the individuals who have reported meaningful data for the variables we are considering in the model, which means that we will filter out individuals who "refused to answer a question," "did not know," etc.

## Primary hypothesis  
As stated above in “Population,” we have received data from about 100,000 individuals among 45,000 households in the United States. 

The primary hypotheses will use a variable indicating usual hours of sleep per day.

The test of our hypothesis will be

$H_0: \beta_1 = \beta_2 = \beta_3 = 0$

$H_1$: At least one $\beta_1 \neq \beta_2 \neq \beta_3 \neq 0$

from the regression model:  
$$ UsualSleep = \beta_0 + \beta_1 ComputerUseFrequency + \beta_2SleepMedication + \beta_3DepressionMedication $$


## Response Variable  
Our response variable measures self-reported sleep. The variable [HRSLEEP](https://nhis.ipums.org/nhis-action/variables/HRSLEEP#codes_section) in IPUMS-NHIS asked participants, "On average, how many hours of sleep do you get in a 24-hour period?" 

Participants were instructed to enter the number of hours slept using whole numbers and rounding up to 30 minutes if necessary.

Answer options: 01 = 1 hour; 02 = 2 hours, ...; 24 = 24 hours; 25 = less than 1 hour, 97 = unknown/refused; 98 = unknown/ascertained; 99 = unknown/refused. We will filter the data by removing those individuals who refused to answer or did not provide a meaningful information.

We will combine responses and create a variable that indicates a small amount of sleep (between 0 and 6 hours) vs. a normal amount of sleep (between 7 and 10 hours) vs. a large amount of sleep (over 10 hours).

We will use multiple regression to model this outcome.

## Explanatory variables

  - Computer Use Frequency

[PCUSEFREQ](https://nhis.ipums.org/nhis-action/variables/PCUSEFREQ#codes_section). 

This will require us to filter out all participants who did not report their computer use (rated as 0 = NIU; 7 = Unknown-refused; 8 = Unknown-not ascertained; 9 = Unknown-don't know) and include only those who answered this question using categorical data (rated as 1 = Never or almost never; 2 = Some days; 3 = Most days; 4 = Every day). We will combine Never/Almost never and Some Days (1 and 2), and Most Days and Every day (3 and 4). 1 and 2: we'll call infrequent use, 3 and 4 - frequent use.

  - Sleep Medication  

[SLEEPMEDS](https://nhis.ipums.org/nhis-action/variables/SLEEPMEDS#codes_section)

The surveyed individuals answer the question *"In the past week, how many times did you take medication to help you fall asleep or stay asleep?"* They answer 0 if they didn't take medication to sleep in the past week; 01-06 for medication taken 1-6 times; 07 for medication taken 7 or more times. 

In the data set, there are many individuals who refused to answer the question, so we will filter out the data leaving only those individuals who answered the question.

  - Depression Medication

[DEPRX](https://nhis.ipums.org/nhis-action/variables/DEPRX#codes_section)

The surveyed individuals answer the question whether they took medication for depression. They answer 0 if they haven't taken any medication for depression and answer 1 if they have taken medication for depression. The variable is a categorical variable so we will transform 1 into "yes" and 0 into "no". We will analyze this in relation to computer use and will observe whether the high frequency computer users have taken depression medication.


#### Potential Confounders  
We suspect that age will be a potential confounder, as sleep habits vary greatly by age. We will also examine race, weight, income, and alcohol use, which we suspect will contribute to sleeping habits.

1. Age [AGE](https://nhis.ipums.org/nhis-action/variables/group?id=demographic_core)  
2. Race [RACENEW](https://nhis.ipums.org/nhis-action/variables/group?id=demographic_ethnicity)  
3. Weight [WEIGHT](https://nhis.ipums.org/nhis-action/variables/WEIGHT#codes_section)  
4. Income [EARNIMP1](https://nhis.ipums.org/nhis-action/variables/group?id=economic_education)  
5. Alcohol Use [ALCDAYSWK](https://nhis.ipums.org/nhis-action/variables/ALCDAYSWK#codes_section)  