---
title: "Draft Proposal"
output:
  html_document:
    code_folding: show

---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE,
                      message = FALSE,
                      options(scipen=999))
```


#### 2020-10-26
### Lika Mikhelashvili, Kelly Ford, Sarah Bingham
**“Hello Statisticians”**

```{r}
#install.packages("ipumsr")
library(ipumsr)
library(ggplot2)
library(moderndive)
ddi <- read_ipums_ddi("nhis_00001.xml.zip")
data <- read_ipums_micro(ddi)
```

Members:

* Lika Mikhelashvili  
* Kelly Ford  
* Sarah Bingham  

## Project Title 
Health Ramifications of Increased Screen Time

## Purpose and Hypotheses
This project focuses on self-reported areas of mental and physical health among 72,831 randomly selected people living in the United States during the year of 2018. These spheres of health include levels of anxiety and depression, whether or not one takes depression medication, hours of sleep per day, weight, and amount of alcohol consumed per year. The goal of this project is to compare these variables to the frequency of computer use in order to more clearly visualize the effect of health on screen time. This project is very relevant to life today where we are depending on technology and computer use more than ever. With this increased screen use however, we may see a decrease in mental and physical health based on limited social interaction and necessary self-isolation from others. Our hope is that this project helps us gain insight into this health and computer usage association within a large, economically and racially diverse sample. 

The **primary hypothesis** is that those with more screen time per week (most days or every day) will have less self-reported usual hours of sleep per day. 

The **secondary hypotheses** are:

1. Individuals with higher frequency of computer use will display more frequent consumption of medication for sleep.
2. Individuals with higher frequency of computer use will display the use of depression medication
3. Individuals with higher frequency of computer use will display higher levels of depression.
## Data  
Our study utilizes data made available online to the public from IPUMS-NHIS (National Health Interview Survey). This data set contains information from 2018, the most recent year available. Our variables include age, race, earnings, weight, alcohol use, sleep habits, effects of depression and anxiety on productivity, depression levels and medications, and frequency of computer use.

## Population  
IPUMS-NHIS data is collected annually of approximately 100,000 individuals throughout 45,000 households in the United States. Each observation represents a separate individual, though this does indicate some individuals cohabitate. Each household has its own serial ID to distinguish which data come from which household. As cohabitation does not violate independence, we will consider all of the data points in our analysis.

Since we are looking specifically at screen time and sleep, our population will be individuals who have self-reported their frequency of computer use [PCUSEFREQ](https://nhis.ipums.org/nhis-action/variables/PCUSEFREQ#codes_section). This will require us to filter out all participants who did not report their computer use (rated as 0, 7, 8, 9) and include only those who answered this question (rated as 1, 2, 3, 4).

## Response Variable  
Our response variable measures self-rated sleep. The variable [HRSLEEP](https://nhis.ipums.org/nhis-action/variables/HRSLEEP#codes_section) in IPUMS-NHIS asked participants, "On average, how many hours of sleep do you get in a 24-hour period?" 

Participants were instructed to enter the number of hours slept using whole numbers and rounding up to 30 minutes if necessary.

We will use linear regression to model this outcome.

## Explanatory Variables  
### Primary hypothesis  
As stated above in “Population”, we have received data from about 100,000 individuals among 45,000 households in the United States. 

The primary hypotheses will use a variable indicating usual hours of sleep her day. We will create this variable from two different sources.

Individuals self-report their usual hours of sleep ([HRSLEEP](https://nhis.ipums.org/nhis-action/variables/HEALTH#codes_section) in NHIS) in response to a question: “About how many hours of sleep do you get per day?”

Answer options: 01= 1 hour; 02 = 2 hours, ...; 24 = 24 hours; 25 = less than 1 hour, 97 = unknown/refused; 98 = unknown/ascertained; 99 = unknown/refused.

We will combine responses and create a variable that indicates a small amount of sleep (between 0 and 6 hours) vs. a normal amount of sleep (between 7 and 10 hours) vs. a large amount of sleep (over 10 hours).

The test of our hypothesis will be

$H_0: \beta_1 = 0$

$H_1: \beta_1 ≠ 0$

from the regression model:  
$$logit(GoodHealth) = \beta_0 + \beta_1 UsualSleep  +\beta'X$$

where $\beta'X$ is a vector of covariates included to control for potential confounding. They are described below.

### Secondary hypotheses  

*1. Individuals with higher frequency of computer use will display more frequent consumption of medication for sleep.*

The variables we'll be using to test the hypothesis are 

  - Sleep Medication - the outcome variable [SLEEPMEDS](https://nhis.ipums.org/nhis-action/variables/SLEEPMEDS#codes_section)

  - Frequency of Computer Use - the explanatory variable [PCUSEFREQ](https://nhis.ipums.org/nhis-action/variables/PCUSEFREQ#codes_section)
  
The surveyed individuals answer the question *"In the past week, how many times did you take medication to help you fall asleep or stay asleep?"* They answer 0 if they didn't take medication to sleep in the past week; 01-06 for medication taken 1-6 times; 07 for medication taken 7 or more times. 

In the dataset there are many individuals who refused to answer the question, so we will filter out the data leaving only those individuals who answered the question.

The test of our hypothesis will be:

- The relationship between frequency of computer use and the amount of sleep medication taken is not significant $H_0: \beta_1 = 0$

- There is a significant relationship between the frequency of computer use and the amount of sleep medication taken $H_1: \beta_1 ≠ 0$

We will try to predict the amount of sleep medication taken based on the frequency of computer use:

$$SleepMedication = \beta_0 + \beta_1ComputerUseFrequency$$

 
*2. Individuals with higher frequency of computer use will display the use of depression medication*

The variables we use to test this hypothesis are: 

  - Depression Medication - the outcome variable [DEPRX](https://nhis.ipums.org/nhis-action/variables/DEPRX#codes_section)

  - Frequency of Computer Use - the explanatory variable [PCUSEFREQ](https://nhis.ipums.org/nhis-action/variables/PCUSEFREQ#codes_section)
  
The surveyed individuals answer the question whether they took medication for depression. They answer 0 if they haven't taken any medication for depression and answer 1 if they have taken medication for depression. The variable is a categorical variable so we will transform 1 into "yes" and 0 into "no". We will analyze this in relation to computer use and will observe whether the high frequency computer users have taken depression medication.

Hypotheses will be: 

  - There is no significant relationship between the frequency of computer use and the use of depression medication $H_0: \beta_1 = 0$

  - There is a significant relationship between the frequency of computer use and the use of depression medication $H_0: \beta_1 ≠ 0$

We will predict the amount of depression medication taken from the frequency of computer use:

$$DepressionMedicationTaken = \beta_0 + \beta_1ComputerUseFrequency$$

*3. Individuals with higher frequency of computer use will display higher levels of depression.*

Explanatory variable used to test the hypothesis: 

  - Level of depression [DEPFEELEVL](https://nhis.ipums.org/nhis-action/variables/DEPFEELEVL#description_section)
  
The surveyed individuals indicate whether their level of depression, the last time they felt depressed, was "a little," "a lot," or "somewhere in between a little and a lot." The indicator variables we will create for these responses are 1- a little, 2 - a lot, 2 - somewhere in between a little and a lot. Some individuals refused to answer the question, so we will filter out the data to include only those individuals that did answer the question.

Hypotheses will be: 

  - There is no significant relationship between the level of depression and the frequency of computer use $H_0: \beta_1 = 0$
  
  - There is a significant relationship between the level of depression and the frequency of computer use $H_0: \beta_1 ≠ 0$

We will try to predict the level of depression based on the frequency of computer use:

$$DepressionLevel = \beta_0 + \beta_1ComputerUseFrequency$$

#### Potential Confounders  
We suspect that age and mental health will be potential confounders, as sleep habits vary greatly by age and mental health. We will also examine race, weight, income, and substance use, which we suspect will contribute to sleeping habits.

1. Age [AGE](https://nhis.ipums.org/nhis-action/variables/group?id=demographic_core)  
2. Mental Health [FLDEPRESMO](https://nhis.ipums.org/nhis-action/variables/FLDEPRESMO#codes_section)  
3. Race [RACENEW](https://nhis.ipums.org/nhis-action/variables/group?id=demographic_ethnicity)  
4. Weight [WEIGHT](https://nhis.ipums.org/nhis-action/variables/WEIGHT#codes_section)  
5. Income [EARNIMP1](https://nhis.ipums.org/nhis-action/variables/group?id=economic_education)  
6. Alcohol Use [ALCDAYSWK](https://nhis.ipums.org/nhis-action/variables/ALCDAYSWK#codes_section)  
7. Sleep Medication [SLEEPMEDS](https://nhis.ipums.org/nhis-action/variables/SLEEPMEDS#codes_section)  
8. Depression Medication [DEPRX](https://nhis.ipums.org/nhis-action/variables/DEPRX#codes_section)  
