---
title: "Draft Proposal"
output:
  html_document:
    code_folding: show

---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE,
                      message = FALSE,
                      options(scipen=999))
```


#### 2020-10-26
### Lika Mikhelashvili, Kelly Ford, Sarah Bingham
**“Hello Statisticians”**

```{r}
library(ipumsr)
library(ggplot2)
library(moderndive)
ddi <- read_ipums_ddi("nhis_00001.xml")
data <- read_ipums_micro(ddi)
```

Members:

* Lika Mikhelashvili  
* Kelly Ford  
* Sarah Bingham  

## Project Title 
Health Ramifications of Increased Screen Time

## Purpose and Hypotheses
This project focuses on self-reported areas of mental and physical health among 72,831 randomly selected people living in the United States during the year of 2018. These spheres of health include levels of anxiety and depression, whether or not one takes depression medication, hours of sleep per day, weight, and amount of alcohol consumed per year. The goal of this project is to compare these variables to the frequency of computer use in order to more clearly visualize the effect of health on screen time. This project is very relevant to life today where we are depending on technology and computer use more than ever. With this increased screen use however, we may see a decrease in mental and physical health based on limited social interaction and necessary self-isolation from others. Our hope is that this project helps us gain insight into this health and computer usage association within a large, economically and racially diverse sample. 

The **primary hypothesis** is that those with more screen time per week (most days or every day) will have less self-reported usual hours of sleep per day. 

## Data  
Our study utilizes data made available online to the public from IPUMS-NHIS (National Health Interview Survey). This data set contains information from 2018, the most recent year available. Our variables include age, race, earnings, weight, alcohol use, sleep habits, depression levels and medications, and frequency of computer use.

## Population  
IPUMS-NHIS data is collected annually of approximately 100,000 individuals throughout 45,000 households in the United States. Each observation represents a separate individual, though this does indicate some individuals cohabitate. Each household has its own serial ID to distinguish which data come from which household. As cohabitation does not violate independence, we will consider all of the data points in our analysis.

Since we are looking specifically at screen time and sleep, our population will be individuals who have self-reported their frequency of computer use [PCUSEFREQ](https://nhis.ipums.org/nhis-action/variables/PCUSEFREQ#codes_section). This will require us to filter out all participants who did not report their computer use (rated as 0, 7, 8, 9) and include only those who answered this question (rated as 1, 2, 3, 4).

## Response Variable  
Our response variable measures self-rated sleep. The variable [HRSLEEP](https://nhis.ipums.org/nhis-action/variables/HRSLEEP#codes_section) in IPUMS-NHIS asked participants, "On average, how many hours of sleep do you get in a 24-hour period?" 

Participants were instructed to enter the number of hours slept using whole numbers and rounding up to 30 minutes if necessary.

We will use linear regression to model this outcome.

## Explanatory Variables  
### Primary hypothesis  
As stated above in “Population”, we have received data from about 100,000 individuals among 45,000 households in the United States. 

The primary hypotheses will use a variable indicating usual hours of sleep her day. We will create this variable from two different sources.

Individuals self-report their usual hours of sleep ([HRSLEEP](https://nhis.ipums.org/nhis-action/variables/HEALTH#codes_section) in NHIS) in response to a question: “About how many hours of sleep do you get per day?”

Answer options: 01= 1 hour; 02 = 2 hours, ...; 24 = 24 hours; 25 = less than 1 hour, 97 = unknown/refused; 98 = unknown/ascertained; 99 = unknown/refused. We will filter the data by removing those individuals who refused to answer or did not provide a meaningful information.

We will combine responses and create a variable that indicates a small amount of sleep (between 0 and 6 hours) vs. a normal amount of sleep (between 7 and 10 hours) vs. a large amount of sleep (over 10 hours).

The test of our hypothesis will be

$ H_0: \beta_1 = 0$

$ H_1: \beta_1 ≠ 0$

from the regression model:  
$$ UsualSleep = \beta_0 + \beta_1 ComputerUseFrequency $$


#### Potential Confounders  
We suspect that age will be a potential confounder, as sleep habits vary greatly by age. We will also examine race, weight, income, and alcohol use, which we suspect will contribute to sleeping habits.

1. Age [AGE](https://nhis.ipums.org/nhis-action/variables/group?id=demographic_core)  
2. Race [RACENEW](https://nhis.ipums.org/nhis-action/variables/group?id=demographic_ethnicity)  
3. Weight [WEIGHT](https://nhis.ipums.org/nhis-action/variables/WEIGHT#codes_section)  
4. Income [EARNIMP1](https://nhis.ipums.org/nhis-action/variables/group?id=economic_education)  
5. Alcohol Use [ALCDAYSWK](https://nhis.ipums.org/nhis-action/variables/ALCDAYSWK#codes_section)  