---
title: "Health Factors Related to Sleep"
author: "Lika Mikhelashvili, Kelly Ford, Sarah Bingham"
date: "12/5/2020"
output:
 html_document:
   code_folding: hide
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE,
                      message = FALSE,
                      options(scipen=999))
```

```{r}
#Packages

library(ipumsr)
library(moderndive)
library(tidyverse)
library(car)
ddi <- read_ipums_ddi("nhis_00001.xml")
data <- read_ipums_micro(ddi)
```


```{r}
#Data wrangling  

hellostats_data <- data %>%
  select(AGE, RACENEW, EARNIMP1, WEIGHT, ALCDAYSWK, HRSLEEP, SLEEPMEDS, DEPRX, PCUSEFREQ) %>%
  filter(ALCDAYSWK %in% c("10", "20", "30", "40", "50", "60", "70", "80"),
         HRSLEEP %in% 1:24,
         SLEEPMEDS %in% c("0", "1", "2", "3", "4", "5", "6", "7"),
         DEPRX %in% c("1", "2"),
         PCUSEFREQ %in% c("1", "2", "3", "4"),
         AGE != 0,
         RACENEW %in% c("10", "20", "30", "40", "50", "60"),
         !(WEIGHT %in% c("996","997", "999"))) %>% 
  mutate(PCUSEFREQ = dplyr::recode(as.integer(PCUSEFREQ), `1` = "Infrequent", `2` = "Infrequent", `3` = "Frequent", `4` = "Frequent"), 
         HRSLEEP = as.integer(HRSLEEP),
         DEPRX = dplyr::recode(as.integer(DEPRX), `1` = "No", `2` = "Yes"),
         SLEEPMEDS =as.integer(SLEEPMEDS),
         ALCDAYSWK = dplyr::recode(as.integer(ALCDAYSWK), `10`=1, `20`=2, `30`=3, `40`=4, `50`=5, `60`=6, `70`=7, `80`=8))
```

## Models  

#### Full model (Sleep medication, depression medication, computer use, weight, alcohol):  
```{r}
full_model <- lm(HRSLEEP ~ SLEEPMEDS + DEPRX + PCUSEFREQ + WEIGHT + ALCDAYSWK, data = hellostats_data)
get_regression_table(full_model)
get_regression_summaries(full_model)
vif(full_model)          
```

#### Reduced model (Sleep medication, depression medication, alcohol):  
```{r}
reduced_model <- lm(HRSLEEP ~ SLEEPMEDS + DEPRX + ALCDAYSWK, data = hellostats_data)
get_regression_summaries(reduced_model)
get_regression_table(reduced_model)
vif(reduced_model)
```

#### Nested F Test:  
```{r}
anova(full_model, reduced_model)
```


# Abstract  

Our project seeks to help understand the environmental conditions that affect an individual's average sleep per night during the Covid-19 pandemic. Our study examines how increased screen time, medication for depression, sleep medication, number of days alcohol is consumed within a week, and weight may affect the average amount of time a person sleeps each night. Individuals stuck at home during the pandemic may be at a greater risk for depression, insomnia, substance use disorder, and of course are more likely to spend a greater amount of time on their computers because of remote work and school.  

Our project uses data collected in 2018 from the National Health Interview Survey (IPUMS-NHIS) to see whether these variables could have an effect on average sleep per night. Both a full and reduced model examine whether these variables have significant effects on the outcome variable of average sleep per night. Though each variable was, in fact, found be a significant predictor of sleep length, we believe our large data set means these findings will have very little practical implication for the general public.  

# Background and Significance  

This project focuses on self-reported areas of mental and physical health among 72,831 randomly selected people living in the United States during the year of 2018. These spheres of health include whether or not one takes depression and/or sleep medication, hours of sleep per day, and amount of alcohol consumed per year. The goal of this project is to compare these variables to amount of sleep in order to more clearly visualize the effect on health. This project is very relevant to life during the pandemic, where we depend on technology and computer use more than ever, factors that we believe effect the quality of sleep. With increased time spent at home, we may see an increase in computer use and a decrease in mental and physical health based on limited in-person social interaction and necessary self-isolation from others. Our hope is that this project gives some insight into the health ramifications associated with sleep within a large sample.  

# Methods  

### Data  

Our study utilizes data made available online to the public from IPUMS-NHIS (National Health Interview Survey). This data set contains information from 2018, the most recent year available. IPUMS-NHIS data is collected annually of approximately 100,000 individuals throughout 45,000 households in the United States. *Our specific data set was collected by obtaining a random sample of 72,831 people living in the U.S who self-reported hours of sleep per night, computer use frequency, weekly alcohol consumption frequency, weight, and how many times within the last week they have taken sleep medication and depression medication.* Each observation represents a separate individual, though the study does indicate some individuals cohabitate. However, each household has its own serial ID to distinguish which data come from which household. As cohabitation does not violate independence, we will consider all of the data points in our analysis. We will analyze the individuals who have reported meaningful data for the variables we are considering in the model, which means that we will filter out individuals who "refused to answer a question," "did not know," etc. *The population that we intend to generalize to is any person living in the United States (roughly about 328 million people).*  
 
### Variables  

The self-reported variables in our analysis include the predictor variable of usual hours slept per day, as well as the following explanatory variables: computer use frequency, weekly alcohol consumption frequency, weight, number of times within the last week they have taken sleep medication, and number of times within the last week they have taken depression medication. Our full model includes all of these explanatory variables, while our reduced model excludes the weight and computer use variable.  

Usual hours of sleep per day is simply the numerical values (1-24) indicating the number of hours that the individual sleeps on an average day. The values that we will filter out are 25 (less than one hour), 97 (refused to answer), 98 (not ascertained), and 99 (unsure).  

Let us look now at our explanatory variables.  

Numerical explanatory variables included sleep medication and weight. In terms of sleep medication, participants answered 0 if they didn't take medication to sleep in the past week, 01-06 for medication taken 1-6 times, and 07 for medication taken 7 or more times. We will filter out answers that are 96 (not applicable), 97 (refused to answer), 98 (not ascertained), and 99 (unsure). Weight tells us the participant's weight in pounds without including clothes or shoes. We will filter out 996 (refused to answer), 997 (not ascertained), and 998 (unknown).

Depression medication was a binary response variable which indicated "2" for taken depression medication in the past week, or "1" for has not taken depression medication in the past week. The variable is a categorical variable so we will transform 2 into "yes" and 1 into "no". We will filter out answers that are 0 (not applicable), 7 (refused to answer), 8 (not ascertained), and 9 (unsure).  

Computer use frequency gave participants the opportunity to give a categorical response of 1 (never or almost never), 2 (some days), 3 (most days), or 4 (every day). We will filter out answers that are 0 (not applicable), 7 (refused to answer), 8 (not ascertained), and 9 (unsure). After receiving this data, we mutated the answers in order to produce a binary response of "frequent" or "infrequent" computer use. We did this by combining responses of 1 and 2 and assigning these to the label "infrequent" and responses 3 and 4 become assigned to "frequent."  

### Analysis

We created two multiple regression models to fit our data. Our first model included weight, sleep medication, and depression medication variables as a function of usual hours slept, so that we could analyze these relationships without accounting for weight or computer use frequency. Our second model included all of the explanatory variables (computer use frequency, alcohol consumption frequency, weight, sleep medication, and depression medication) as a function of usual hours slept. Because our first model is "nested" within our second model, we will refer to this as the "reduced model". Our second model in this case would be the "full model". 

For both of these models, we calculated the p-values of each of the variables in order to determine the significance in explaining usual hours slept. We then ran a nested F test in order to determine whether we needed the full model to explain the data, or if the reduced model did the job itself. 

# Results  

```{r}
sleep_mean <- mean(hellostats_data$HRSLEEP)
```

```{r}
weight_mean <- mean(hellostats_data$WEIGHT)
```

We began with a large national data set of 72,831 people. After filtering for non-applicable responses, omitting variables with more than two categories (of which we were not yet capable of analyzing statistically), and computer use frequency data, 11,672 individuals remained. Average amount of sleep per night is 7.08 hours with an average body weight of 177.49 pounds. The data consisted of overwhelmingly White participants, with 84.21% of individuals self-identifying as White, and 9.07% as Black or African-American, 3.49% as Asian, 1.15% Alaskan Native/American Indian, 2.08% as multiple race, and 0% as other. It should, therefore, be noted that this data set does not adequately represent individuals who do not identify as White and should not be generalized as representing all races adequately.  

Our group analyzed two separate models, full and reduced; however, the coefficients for **every** variable, regardless of model, was found to be significant ($p < .05$). Though all of the coefficients were significant predictors of sleep duration, both the reduced ($R^2 = .007$) and the full models ($R^2 = .011$) had very small $R^2$ numbers, indicating that the reduced model only accounted for .7% and the full model only 1.1% of variability. After picking a reduced model that omitted weight and computer use frequency, we ran a nested F-test to determine whether the full model was necessary since all the coefficients had very small P values. Our significant F-test determined the full model was a stronger model and that the two extra variables should be retained ($F < .05$, $p < .05$).  

Finally, we believe the extremely small p-values mean that though all our coefficients were technically significant predictors of the average amount of sleep per night, it is in part due to the extremely large data set ($N = 11,672$). Despite the significance of these predictors, we believe the actual practical application of these findings needs a more specialized study. With such a large data set, it is likely that the significance of sleep time could so small as to be inconsequential; i.e., an extra 10 minutes of sleep per night may be statistically significant in our model, but it is inconsequential to the individual's overall lived experience.  Additionally, the more frequently alcohol is consumed throughout the week is significantly positively correlated to an increase in average time slept per night, but this study fails to account for the *quality* of said sleep.  

# Discussion  

The main goal of our analysis is to assess the relationship between sleep duration in relation to computer use, sleep medication, depression medication, alcohol, and weight. Based on the full model, we can infer the relative effect of each variable on the sleep duration. One additional sleep medication (variable `SLEEPMEDS`) taken in one week decreases the hours slept by 0.03 hour. When the individual has taken a depression medication (`DEPRXYes`), the hours of sleep increases by 0.274 hours. An infrequent computer use increases the hours slept by 0.184, so an individual with high frequency of computer usage has a relatively shorter sleep duration. One unit increase in number of alcohol drinks consumed in a week (`ALCDAYSWK`) increases the hours slept, while one unit increase in individual's weight (`WEIGHT`) decreases the sleep duration. 

Our findings confirm our initial hypothesis that a more frequent computer usage is correlated to a decreased sleep duration. This is a relevant finding to the current pandemic since the drastic increase computer and smart-phone usage has supposedly led to an unstable sleep schedule for many. However, based on the data and the tests we conducted, we cannot claim a causal relationship between the two factors. 

When accounting for other factors in the model, the relationship between hours slept (`HRSLEEP`) and the number of alcohol drinks per week (`ALCDAYSWK`) shows a slight positive trend, based on the figures 1.1 and 1.2. This relationship is evident in both infrequent and frequent computer use cases and regardless of the depression medication intake.

**Figure 1.1**

```{r}
ggplot(hellostats_data, aes(x=ALCDAYSWK, y=HRSLEEP, color=SLEEPMEDS)) +
  geom_point() +
  geom_smooth(method = lm, se=FALSE, formula=y~x) +
  facet_grid(PCUSEFREQ ~ DEPRX)
```

**Figure 1.2**

```{r}
ggplot(hellostats_data, aes(x=ALCDAYSWK, y=HRSLEEP, color=SLEEPMEDS)) +
  geom_point() +
  geom_smooth(method = lm, se=FALSE, formula=y~x) +
  facet_wrap(~DEPRX)
```


### Limitations  

The limitation of our analysis is the fact that the values in all variables are self-reported, which makes the data less accurate. Also, we were limited to IPUMS variables and the survey texts they used. For example, in case of `PCUSEFREQ` variable, instead of having the variable be categorical, it would be much more accurate to have numeric variable with the number of hours of computer usage. Due to less flexibility in the available variables, we cannot infer any causal relationship between the hours slept and computer/health medication usage. The lack of racial diversity in this sample also fails to represent the experiences of BIPOC and people of the global majority. More sampling should be conducted which adequately represents the experiences of racial groups besides those who identify as White.

### Conclusion  

Our findings were statistically significant, yet inconclusive, and we therefore cannot claim a causal relationship between any of the explanatory variables and sleep duration. The sample data is so big that slight changes in explanatory variables result in very slight changes in the outcome variables. What's more, the explanatory variables describe a very small proportion of variability in sleep duration (R^2 is small). However, our analysis shows that frequent computer usage is associated with a decreased sleep length and alcohol usage is correlated with increased sleep length. 

